Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-15 Ben_Manuscripts/structure_paper/Draft.tex
Raw WAXS integration

  Our implementation requires long simulation times to achieve high degrees
  of crosslinking. For that reason we did not crosslink all systems tested, but only the most
  promising structure. We show that crosslinking does not significantly 
  change any of our drawn conclusions in Section ~\ref{section:xlink}.

@@

  We explored the pore composition by measuring the average number densities of
  various monomer components as a function of distance from the pore centers. The
  centers of the pores were located using the same method used to calculate the
  equilibrium pore spacing. We looked at the average number density of sodium
  ions, aromatic rings and carbon atoms making up the monomer tails. The radial
  distance of all atoms in each group from the pore centers are binned, then
  normalized by the volume of the annulus defined by the bin edges and z box
  vector.
  %BJC: Maybe write an equation here? It's hard to say precisely how I did it
  %in words without it sounding confusing

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation relates the DC ionic conductivity to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the boltzmann constant,
  $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$. 
  Sodium ion diffusion coefficients were found by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the einstein relation \cite{einstein_investigations_1956}.] We
  looked at the MSD plot to determine where to begin and end a linear fit. Ion
  concentration was measured with respect to the volume of the entire unit cell.

  The second method, termed the 'Collective Diffusion' model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$ of the system can be
  calculated as: $$ \gamma = \dfrac{D_Q}{k_b T} $$. Conversion to ionic
  conductivity is achieved by multiplying by channel length and dividing by the
  membrane cross sectional area.  $D_Q$ is the diffusion coefficient of the
  collective variable Q. It can be calculated using the einstein relation.  A
  full derivation of the model can be accessed elsewhere 
  \cite{liu_collective_2013}.

@@

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equlibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing (\ref{point:monomernum}), we ran simulations of
  systems created with 4 - 8 monomers per layer. Systems were built in both the
  parallel displaced and sandwiched configurations with layers initially spaced
  3.7 \AA~ apart. Equilibrated configurations were prepared according to the dry
  equilibration procedure. All systems are stable after 400 ns of simulation. In
  a sense, all systems are at least metastable, addressing
  (\ref{point:metastable}), however not all will make physical sense or fit the
  experimental profile that we are looking to match. Figure~\ref{fig:p2p} shows
  the pore spacing for all systems tested. Systems built with 5 monomers in each
  layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  We learned that layers are well-defined and persistent, answering
  (\ref{point:layers}). We established our conclusion by plotting the pair
  correlation function, $g(z)$ calculated between atoms along the length of the
  pores (Fig.~\ref{fig:zdf}). We measured $g(z)$ with respect to aromatic rings in
  the head groups and, separately, with respect to carbon atoms in the alkane
  chains. Using a fourier transform of $g(z)$, we see that sandwiched
  configuration layers stack 4.39 \angstrom apart while parallel displaced
  configuration layers stack 4.38 \angstrom apart.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{The pore spacing (given in nm) of the model increases as
	number of monomers in each layer increases. The pore spacing of a system
	starting in the sandwiched configuration is systematically lower than 
	that started in an offset configuration. Systems built with 5 monomers
	per layer in a parallel displaced configuration result in a pore spacing
	closest to the experimental value of 4.12 nm}~\label{fig:p2p}
  \end{figure}

@@

  We answer question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. Simulated patterns were generated using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure ~\ref{fig:XRDsim}.

  Simulated XRD of the sandwiched configuration contains all experimental
  features except for R-helix. R-alkanes, R-spots and R-pores appear close to their
  expected locations. R-$\pi$ is also present, however it intersects R-alkanes at
  a q value lower than experiment meaning the head groups in our model prefer 
  to stack farther apart.

  The parallel displaced configuration results in a simulated XRD pattern with
  the closest match to experiment. It produces the only pattern that exhibits
  all major reflections. R-alkanes, R-pores and R-$\pi$ appear as they do in the
  sandwiched configuration. R-spots appears, however with a lower intensity
  relative to R-alkanes when compared to the sandwiched configuration. R-helix
  appears due to the parallel displaced aromatic rings. It is a subharmonic
  of R-$\pi$ since the nearest vertically stacked head group to any given head
  group is 7.4 \AA away.

@@

  The peaks in the tilt angle distribution are consistent with the location of
  R-spots. The peaks of interest in Figures \ref{fig:offset_tails} and
  \ref{fig:layered_tails} are located at $\pm$ 33 $\degree$ which is the same
  location where the highest intensity of spots are located on the simulated
  patterns. We confirmed this conclusion by radially integrating the 2D WAXS
  pattern for $\left|\mathbf{q}\right|$ values between 1.57 and 1.4 (real space between 4 and 4.5 \AA). We
  observe that distinct peaks appear ca. 30 $\degree$, in close agreement with
  the previously measured angle distribution
  (Figs.~\ref{fig:offset_integration}~and~\ref{fig:layered_integration}). We performed the
  same integration on the raw experimental data and found the angle at which
  R-spots reaches its highest intensity to be $\pm$ 37 $\degree$ which is a
  reconcilable difference with our simulated results.

@@

  \begin{figure}[!htb]
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:layered_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure}
%MRS: 
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
  \end{figure}

@@

%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom as suggested by experiment.
  When systems are built with layers stacked 5.0 \AA~ apart then
  equilibrated, we observe long-term stability of a qualitatively different
  configuration suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). We studied this type of
  system in both the parallel displaced and sandwiched configurations.

@@

  In order to address (\ref{point:poredefinition}), we plotted the number of
  densities of heavy atoms in the head group, carbon atoms in the tail region
  and the sodium ions (Figure~\ref{fig:densities}). For the head group region,
  we used the carbon atoms making up the aromatic ring. For the tail region we
  used only carbon atoms of the monomer tails (See Supplemental Information for
  diagram). Histograms are average over at least 50 ns of equilibrated trajectory.

@@

  The sandwiched configuration (Fig.~\ref{fig:layered_density}) shows a relatively
  ordered pore structure. The number densities of components in all other
  configurations are comparable. In contrast, sodium ions and head groups in the
  sandwiched configuration are highly concentrated away from the pore center
  meaning that there is vacancy in the pore region. A vacant pore is very likely
  to be unstable in a real system, further supporting the physical realism of the
  parallel displaced configuration.

@@

  In systems built with 5 wt~\% water, the pore region becomes filled with
  water. We plotted the number density of components in this system and see that
  the pores become a soup of water molecules and sodium ions (Fig.
  \ref{fig:water_density}. This system is a closer representation of the
  H\textsubscript{II} phase which is typically synthesized with ca. 8 wt \%
  water. We believe that the hydrated membrane pores will be valuable for aqueous
  separations. Further analysis of and transport within the H\textsubscript{II}
  phase are left for future studies.

@@

  The calculated values of ionic conductivity are higher than experiment by an
  order of magnitude. One can justify the reason for this result by considering
  the real system studied experimentally. The ionic conductivity measurement to
  which we are comparing was done on a \SI{80}{\micro\metre} thick film, nearly
  10,000 times thicker than our simulated system. The thick film likely has
  defects leading to non-contiguous pores and imperfect alignment.  It has been
  shown that there is a large dependence of ionic conductivity on the alignment
  of the pores. The ionic conductivity of an unaligned film is ca. 85 times lower
  than that of a nearly aligned film referenced here. We hypothesize that a thin,
  perfectly aligned film would have a value of ionic conductivity in agreement
  with our model.

  \begin{figure}
        \centering
        \includegraphics[width=0.5\linewidth]{IC_offset.png}
        \caption{Calculated ionic conductivity using the Nernst-Einstein relation
        and Collective Diffusion model agree with error. Both methods give calculated
        values of ionic conductivity an order of magnitude higher than the experimental
        value}
        \label{fig:conductivity}
  \end{figure}

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the crosslinking algorithm to the equilibrated
  parallel-displaced configuration built with 5 monomers per layer. We simulated
  the crosslinked system in the NPT ensemble for 100 ns. The distance between
  pores shrinks by 0.4 \angstrom~ after the system is crosslinked. All major
  features are still present in the X-ray diffraction, however at lower
  intensities (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using
  the Nernst-Einstein relation and found that it is lower in the crosslinked
  system (Fig.~\ref{fig:IC_xlink}).

@@

  % add stuff about pore structure and 'disordered' basin once there are concrete results 
  We have used a detailed molecular model of the Col\textsubscript{h} phase
  formed by NA-GA3C11 in order to study its nanoscopic structure. While there
  have been efforts to model formation of various liquid crystalline phases with
  molecular dynamics, to our knowledge there have been no studies which attempt
  to examine their structure with the same level of detail presented here.
  Evidence strongly supports that layers are composed of 5 monomers. We have
  confirmed that monomers stay partitioned into layers. We have explored the
  affect of two different $\pi$-$\pi$ stacking modes on the equilibrated membrane
  strucure. Simulated diffration patterns generated from MD trajectories suggest
  that the parallel-displaced configuration produces a structure with the closest
  match to experiment. Finally, water is not needed to create well-definied pore
  structures. Now that we have a good idea of what the membrane structure should
  look like, we will evaluate transport of various solutes within the system. We
  will apply the knowledge gained from this study in order to suggest
  improvements to the existing system as well as to evaluate new unsynthesized
  LLC systems.

>>>>

  Our implementation requires long simulation times to achieve high degrees
  of crosslinking. For that reason we did not crosslink all systems tested, but only the most
  promising structure. We show that crosslinking does not significantly 
  change any of our drawn conclusions in Section 3.6.

@@

  We explored the pore composition by measuring the average number densities of
  various monomer components as a function of distance from the pore centers.  We
  looked at the average number density of sodium ions, aromatic rings and carbon
  atoms making up the monomer tails. The radial distance of all atoms in each
  group from the pore centers are binned, then normalized by the volume of the
  annulus defined by the bin edges and z box vector.
  %BJC: Maybe write an equation here? It's hard to say precisely how I did it
  %in words without it sounding confusing

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation relates the DC ionic conductivity to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the boltzmann constant,
  $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$ 
  Sodium ion diffusion coefficients were found by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. Ion
  concentration was measured with respect to the volume of the entire unit cell.

  The second method, termed the Collective Diffusion model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$ of the system can be
  calculated as: $$ \gamma = \dfrac{D_Q}{k_b T} $$ Conversion to ionic
  conductivity is achieved by multiplying by channel length and dividing by the
  membrane cross sectional area.  $D_Q$ is the diffusion coefficient of the
  collective variable Q. It can be calculated using the einstein relation.  A
  full derivation of the model can be accessed elsewhere 
  \cite{liu_collective_2013}.

@@

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing (\ref{point:monomernum}), we ran simulations of
  systems created with 4 - 8 monomers per layer. Systems were built in both the
  parallel displaced and sandwiched configurations with layers initially spaced
  3.7 \AA~apart. Equilibrated configurations were prepared according to the dry
  equilibration procedure. All systems are stable after 400 ns of simulation. In
  a sense, all systems are at least metastable, addressing
  (\ref{point:metastable}), however not all will make physical sense or fit the
  experimental profile that we are looking to match. Figure~\ref{fig:p2p} shows
  the pore spacing for all systems tested. Systems built with 5 monomers in each
  layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{Systems built with 5 monomers per layer in a parallel
		displaced configuration result in a pore spacing closest to the experimental
		value of 4.12 nm. The pore spacing of the model increases as number of monomers
		in each layer increases. The pore spacing of a system starting in the
		sandwiched configuration is systematically lower than that started in an offset
		configuration. }~\label{fig:p2p}
  \end{figure}

  We learned that layers are well-defined and persistent, answering
  (\ref{point:layers}). We established our conclusion by plotting the pair
  correlation function, $g(z)$, calculated between atoms along the length of the
  pores (Fig.~\ref{fig:zdf}). We measured $g(z)$ with respect to aromatic rings in
  the head groups and, separately, with respect to carbon atoms in the alkane
  chains. Using a fourier transform of $g(z)$, we see that sandwiched
  configuration layers stack 4.39 \angstrom~ apart while parallel displaced
  configuration layers stack 4.38 \angstrom~ apart.

@@

  We answer question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. Simulated patterns were generated using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure \ref{fig:XRDsim}.

  Simulated XRD of the sandwiched configuration contains all experimental
  features except for R-helix. R-alkanes, R-spots and R-pores appear close to their
  expected locations. R-$\pi$ is also present, however it intersects R-alkanes at
  a $q_z$ value lower than experiment meaning the head groups in our model prefer 
  to stack farther apart.

  The parallel displaced configuration results in a simulated XRD pattern with
  the closest match to experiment. It produces the only pattern that exhibits
  all major reflections. R-alkanes, R-pores and R-$\pi$ appear as they do in the
  sandwiched configuration. R-spots appears, however with a lower intensity
  relative to R-alkanes when compared to the sandwiched configuration. R-helix
  appears due to the parallel displaced aromatic rings. It is a subharmonic
  of R-$\pi$ since the nearest vertically stacked head group to any given head
  group is 7.4 \AA~ away.

@@

  The peaks in the nearest neighbor angle distribution are consistent with the
  location of R-spots. The peaks of interest in Figures \ref{fig:offset_tails}
  and \ref{fig:layered_tails} are located at $\pm$ 33 $\degree$ which is the same
  location where the highest intensity of spots are located on the simulated
  patterns. We confirmed this conclusion by radially integrating the 2D WAXS
  pattern for $\left|\mathbf{q}\right|$ values between 1.4 and 1.57 (between 4
  and 4.5 \AA~ in real space). We observe that distinct peaks appear ca. 30
  $\degree$, in close agreement with the previously measured angle distribution
  (Figs.~\ref{fig:offset_integration}~and~\ref{fig:layered_integration}). We
  performed the same integration on the raw experimental data and found the angle
  at which R-spots reaches its highest intensity to be $\pm$ 37 $\degree$ which
  is a reconcilable difference with our simulated results.

@@

  \begin{figure}[!htb]
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:layered_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure} 
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
  \end{figure}

@@

%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom as suggested by experiment.
  When systems are built with layers stacked 5.0 \AA~apart then
  equilibrated, we observe long-term stability of a qualitatively different
  configuration suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). We studied this type of
  system in both the parallel displaced and sandwiched configurations.

@@

  In order to address (\ref{point:poredefinition}), we plotted the number of
  densities of heavy atoms in the head group, carbon atoms in the tail region
  and the sodium ions (Figure~\ref{fig:densities}). For the head group region,
  we used the carbon atoms making up the aromatic ring. For the tail region we
  used only carbon atoms of the monomer tails (See Supplemental Information for
  diagram). Histograms are averaged over at least 50 ns of equilibrated trajectory.

@@

  The sandwiched configuration (Fig. \ref{fig:layered_density}) shows a relatively
  ordered pore structure. The number densities of components in all other
  configurations are comparable. In contrast, sodium ions and head groups in the
  sandwiched configuration are highly concentrated away from the pore center
  meaning that there is vacancy in the pore region. A vacant pore is very likely
  to be unstable in a real system, further supporting the physical realism of the
  parallel displaced configuration.

@@

  In systems built with 5 wt~\% water, the pore region becomes filled with
  water. We plotted the number density of components in this system and see that
  the pores become a soup of water molecules and sodium ions (Fig.
  \ref{fig:water_density}). This system is a closer representation of the
  H\textsubscript{II} phase which is typically synthesized with ca. 8 wt \%
  water. We believe that the hydrated membrane pores will be valuable for aqueous
  separations. Further analysis of and transport within the H\textsubscript{II}
  phase are left for future studies.

@@

  The calculated values of ionic conductivity are higher than experiment by an
  order of magnitude. One can justify the reason for this result by considering
  the real system studied experimentally. The ionic conductivity measurement to
  which we are comparing was done on a \SI{80}{\micro\metre} thick film, nearly
  10,000 times thicker than our simulated system. The thick film likely has
  defects leading to non-contiguous pores and imperfect alignment.  It has been
  shown that there is a large dependence of ionic conductivity on the alignment
  of the pores. The ionic conductivity of an unaligned film is ca. 85 times lower
  than that of a nearly aligned film referenced here. We hypothesize that a thin,
  perfectly aligned film would have a value of ionic conductivity in close 
  agreement with our model.

  \begin{figure}
        \centering
        \includegraphics[width=0.5\linewidth]{IC_offset.png}
        \caption{Calculated ionic conductivity using the Nernst-Einstein relation
        and Collective Diffusion model agree with error. Both methods give calculated
        values of ionic conductivity an order of magnitude higher than the experimental
        value.}
        \label{fig:conductivity}
  \end{figure}

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the crosslinking algorithm to the equilibrated
  parallel-displaced configuration built with 5 monomers per layer. We simulated
  the crosslinked system in the NPT ensemble for 100 ns. The distance between
  pores shrinks by 0.4 \angstrom~ and the distance between layers increases by
  0.04 \AA~ after the system is crosslinked. All major features are still present
  in the simulated XRD patterns, however at lower intensities
  (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using the
  Nernst-Einstein relation and found that it is lower in the crosslinked system
  (Fig.~\ref{fig:IC_xlink}).

@@

  % add stuff about pore structure and 'disordered' basin once there are concrete results 
  We have used a detailed molecular model of the Col\textsubscript{h} phase
  formed by Na-GA3C11 in order to study its nanoscopic structure. While there
  have been efforts to model formation of various liquid crystalline phases with
  molecular dynamics, to our knowledge there have been no studies which attempt
  to examine their structure with the same level of detail presented here.
  Evidence strongly supports that monomers stay partitioned into layers and that
  each layer contains 5 monomers. We have explored the affect of two different
  $\pi$-$\pi$ stacking modes on the equilibrated membrane strucure. Simulated
  diffration patterns generated from MD trajectories suggest that the
  parallel-displaced configuration produces a structure with the closest match to
  experiment. Finally, water is not needed to create well-definied pore
  structures.

  Now that we have a good idea of what the membrane structure should
  look like, we will evaluate transport of various solutes within the system. We
  will apply the knowledge gained from this study in order to suggest
  improvements to the existing system as well as to evaluate new unsynthesized
  LLC systems.
